Haploid Selection, Sex Ratio Bias, and Transitions Between Sex-Determination Systems

### 2018-04-02 LaTeX/main_revised.tex
refining responses, attempting to pass off to mscott

\end{flushleft}
% Please keep the abstract below 300 words
%currently 312 words
\section*{Abstract}
Sex determination is remarkably dynamic; many taxa display shifts in the location of sex-determining loci or the evolution of entirely new sex-determining systems. 
Predominant theories for why we observe such transitions generally conclude that novel sex-determining systems are favoured by selection if they equalise the sex ratio or increase linkage with a \textcolor{red}{sexually-antagonistic locus}. 
We use population genetic models to extend these theories in two ways: 
(1) We explicitly consider selection on loci very tightly linked to the ancestral sex-determining loci, e.g., within the non-recombining region of the ancestral sex chromosomes. 
Variation at such loci can favour the spread of new sex-determining systems in which the heterogametic sex changes (XY to ZW or ZW to XY) and the new sex-determining region is less closely linked (or unlinked) to the locus under selection, which is not predicted by previous theory. 
(2) We consider selection upon haploid genotypes either during gametic competition (e.g., pollen/sperm competition involving multiple males) or meiosis (i.e., non-Mendelian segregation), which can cause the zygotic sex ratio to become biased. 
We find that associations with haploid selected loci can drive transitions between sex-determining systems \textcolor{red}{, without requiring sexually-antagonistic selection in diploids}. 
Unexpectedly, with haploid selection, transitions between male and female heterogamety can also evolve where linkage with the sex-determining locus is weakened. 
Furthermore, sex-ratio biases may increase or decrease with the spread of new sex chromosomes. 
Thus, we find that transitions between sex-determining systems cannot be simply predicted by selection to equalise the sex ratio. 
Overall, our models reveal that transitions between sex-determining systems, particularly transitions where the heterogametic sex changes, can be driven by loci in previously unpredicted genomic locations that experience selection during diploid and/or haploid phases.
These results predict conditions under which sex-determining systems are likely to be labile and draw connections with sex-ratio evolution.

@@

%Tight linkage Model Feature and Result
One novel feature of the models developed here is that we explicitly \textcolor{red}{(is this wording fair to vD\&K?)} consider loci in very tight linkage with the ancestral sex-determining locus (e.g., within the non-recombining region of a sex chromosome). % in the results section: `Tight linkage with the ancestral sex-determining locus'). 
Our analysis shows that, under various forms of selection, the ancestral-X can carry suboptimal alleles for females and/or the ancestral-Y can carry female-beneficial alleles. 
This, in turn, can favour a new ZW sex-determining locus that has weaker linkage with loci under selection, which has not been predicted by previous theory. 
That is, very tight linkage with the ancestral GSD locus can favour trans-GSD transitions in which associations with selected loci are actually weakened.

@@

Here, we assume that the $\mathbf{M}$ locus is `epistatically dominant' over the $\mathbf{X}$ locus such that zygotes with at least one $m$ allele develop as females with probability $k$ and as males with probability $1-k$, regardless of the $\mathbf{X}$ locus genotype. 
With $k=0$, the $m$ allele is a masculinizer (a neo-Y allele) and with $k=1$ the $m$ allele is a feminizer (a neo-W allele). 
With intermediate $k$, we can interpret $m$ as an environmental sex-determination (ESD) allele, such that zygotes develop as females in a proportion ($k$) of the environments they experience. 
The assumption that derived sex-determining loci are epistatically dominant is motivated by empirical systems in which multiple sex determining alleles segregate (i.e., X, Y, Z, and W alleles present), such as, cichlid fish~\cite{Ser:2010iq}, platyfish (\textit{Xiphophorus maculatus}~\cite{Kallman1968}), houseflies (\textit{Musca domestica} \textcolor{red}{CITATION}), western clawed frogs (\textit{Xenopus tropicalis}~\cite{Roco2015}) and \textit{Rana rugosa}~\cite{Ogata:2007jm}.  
Nevertheless, our supplementary analysis file (\nameref{file:Mathematica}) allows other dominance relationships between loci to be specified. 
%We also analyze a model of maternally-controlled environmental sex-determination, where mothers with at least one $m$ allele produce daughters with probability $k$.

@@

\noindent\textit{Conclusion 1A:}
\textbf{selection on loci within the non-recombining region around the sex-determining locus ($r\approx0$) allows the invasion of less closely linked sex-determining alleles ($r<R\leq1/2$) during a trans-GSD transition}.
%\textcolor{red}{[new result for sex-antagonism and overdominance, but already shown by Kozielska for ploidy-antagonism]}
\vspace{0.5cm}

\noindent This does not depend on the form of selection maintaining a polymorphism (sexually-antagonistic selection, overdominance, ploidally-antagonistic selection, or some combination, Fig \ref{fig:SexAntagTighter}, \nameref{fig:temporalOverdominance}, \nameref{fig:regionPloidAntag}, and \nameref{fig:SexAntagTighterMaleDrive}).  
The conditions become more restrictive, however, with increasing recombination ($R$) between the new sex-determining locus and the selected locus.  
The invasion of completely unlinked neo-W alleles ($R = 1/2$) can occur with overdominance in males or with haploid selection (as shown for the case of male drive opposed by diploid selection in \cite{Kozielska:2010vm}) but is not possible with only sexually-antagonistic selection if selection is directional in each diploid sex (see \nameref{file:Mathematica}).
%\vspace{0.5cm}

@@

However, with haploid selection and $r<1/2$, the additional term in Eq \eqref{eq:lambda_neoW} can be positive.
This implies, 
\vspace{0.5cm}

\noindent\textit{Conclusion 1B:}\
\textbf{when there is an intermediate rate of recombination between the ancestral sex-determining allele and a selected locus ($0<r<1/2$), haploid selection allows less closely linked sex-determining alleles ($r<R$) to invade during trans-GSD transitions ($\lambda_{W'}^{(XY)}>1$).}
%\textcolor{red}{[new result because Kozielska had ancestrally-tight sex-linkage and Ubeda had ESD]}
\vspace{0.5cm}

\noindent\textcolor{red}{\textit{Conclusion 2A:}\
\textbf{New sex-determining alleles (causing cis- \textit{or} trans-GSD transitions) can spread if the arise in closer linkage with a locus that experiences haploid selection than the ancestral-sex-determining locus ($R<r$).} (This is different but similar to Ubeda et al (2015), who started from ESD)}

\noindent\textcolor{red}{\textit{Conclusion 2B:}\
\textbf{New sex-determining alleles (causing trans-GSD transitions) can spread if they arise in looser linkage with a locus that experiences haploid selection than the ancestral-sex-determining locus ($r<R$). } }

%These transitions are unusual because, when $R>r$, associations that selection has built up between alleles more favourable in one sex and alleles that determine sex will be weakened. 
%Mean diploid fitness therefore decreases during heterogametic transitions that create looser sex-linkage (Fig \ref{fig:Combination_MeanFit}B,D). 
%Eq \eqref{eq:lambda_neoW} shows that, with weak selection, neo-W alleles can invade an XY system for a large number of selective regimes. 
%To clarify the parameter space under which $\lambda_{W'}^{(XY)}>1$ , we consider several special cases. 
%Firstly, if the $\mathbf{A}$ locus is unlinked to the ancestral sex-determining region ($r=1/2$), a more closely linked neo-W ($R<1/2$) can always invade because there is no ancestral association between $A$ alleles and sex chromosomes in males, $\left( \hat{p}^\male_Y-\hat{p}^\male_X \right)=0$, see Eq \eqref{eq:freq_diffs}. 
%The second term in Eq \eqref{eq:lambda_neoW} therefore disappears and invasion depends only on the sign of $(r-R)$, as in the case of the neo-Y. 
%Indeed, invasion typically occurs when the neo-W is more closely linked to the selected locus than the ancestral sex-determining region (Fig \ref{fig:Combination_Centimorgans}).
To clarify the parameter space under which invasion occurs despite looser sex-linkage ($\lambda_{W'}^{(XY)}>1$ despite $R>r$), we focus on the special case where $R=1/2$ and $r<1/2$ (e.g., the selected locus is on the ancestral sex chromosome and the novel sex-determining locus arises on an autosome). 
In Table \ref{tab:specialcases} we give the conditions where invasion occurs when we further assume that haploid selection only occurs in one sex (e.g., during male meiosis only) and dominance coefficients are equal in the two sexes, $h^\female=h^\male$. 
These special cases indicate that neo-W invasion occurs for a large fraction of the parameter space, even if the neo-W uncouples the sex-determining locus from a locus under selection. 
We show that neo-W alleles can spread for when they are more closely \textit{or} more loosely linked to a locus that experiences haploid selection in Fig \ref{fig:Combination_Centimorgans} (c.f., Fig \ref{fig:SexAntagTighter}A for diploid sexually-antagonistic selection alone). 
In Fig \ref{fig:SexRatioBad}, we also show that trans-GSD transitions in sex determination occur during which linkage with a haploid selected locus increases \textit{or} decreases.

@@

Previous research has shown that when the ancestral sex-determining locus is linked to a locus that experiences haploid selection (e.g., meiotic drive), alleles at a new, unlinked sex-determining locus can invade and restore equal sex ratios~\cite{Kozielska:2010vm}. 
Consider, for example, the case where the $\mathbf{A}$ locus is linked to the ancestral sex-determining locus ($r<1/2$) and experiences meiotic drive in males only ($\alpha^\male_\Delta \neq 0$, $\alpha^\female_\Delta=0$), without gametic competition ($t^\female=t^\male=0$).
In this case, the zygotic sex ratio can be initially biased only if the ancestral sex-determining system is XY (Fig \ref{fig:model_outline}A and Fig \ref{fig:SexRatioBad}B). 
%If the ancxestral sex-determining system is ZW, the zygotic sex ratio will be 1:1 because diploid sex is determined by the proportion of Z-bearing versus W-bearing eggs and meiosis in females is fair (Fig \ref{fig:Combination_Turnover}D).
If Fisherian sex-ratio selection were dominant, we would thus expect a difference in the potential for XY to ZW and ZW to XY transitions. 
%However, to leading order with selection weak relative to recombination, we find that sex-ratio selection favours the spread of a neo-W by an amount equl to the spread of a neo-Y into a ZW system (relabeling XY as the heterogametic sex and reversing the labels of males and females in Table \ref{tab:specialcases})
However, invasion by a neo-W allele into an XY system and invasion by a neo-Y allele into a ZW system occur under the same conditions ($\lambda_{Y'}^{(XY)} = \lambda_{W'}^{(ZW)}$ and $\lambda_{W'}^{(XY)}=\lambda_{Y'}^{(ZW)}$, at least to order $\epsilon^2$), implying that
\vspace{0.5cm}

@@

\noindent\textcolor{red}{\textit{Conclusion 3:}
\textbf{when selection is weak relative to recombination, sex-ratio selection is exactly equal in magnitude to the selection imposed by associations with selected alleles. } }
%\textcolor{red}{[new result]}
\vspace{0.5cm}

@@

We next consider the case where the new sex-determining mutation, $m$, causes sex to be determined probabilistically or by heterogeneous environmental conditions (environmental sex determination, ESD), with individuals carrying allele $m$ developing as females with probability $k$.
\textcolor{red}{In our deterministic model this means the fraction female is exactly $k$, even when $m$ is rare (exploring the effect of the variance induced by ESD would also be interesting).}
Here, we do not assume that the environmental conditions that determine sex also differentially affect the fitness of males versus females. 
Such correlations can favour environmental sex-determining systems by allowing each sex to be produced in the environment in which it has highest fitness; in the absence of these correlations, previous theory would predict that ESD is favoured when it produces more equal sex ratios than the ancestral system (see reviews by~\cite{Charnov:1982wg,Bull:1983vi,West:2009we}). %(e.g., maternal condition, mate quality, age, or host size)

@@

\noindent\textit{Conclusion 3:}
\textbf{
transitions from genetic to environmental sex-determination are not straightforwardly predicted by selection to balance the zygotic sex ratio when haploid selection is present. 
}
%\textcolor{red}{[new result]}

@@

%HAPLOID SELECTION (genomic location of haploid selected loci and comparative patterns) 
We have shown that the spread of new sex-determining systems can be driven by loci experiencing haploid selection. 
%Our results suggest that haploid selected loci that have been involved in driving turnovers between sex-determining systems could occur around the ancestral- or novel-sex-determining regions. 
%Our results suggest that polymorphic loci that are ancestrally sex-linked and under sex-specific selection could also drive heterogametic transitions between sex-determining systems. 
%As with sexually-antagonistic selection, the presence of haploid selected loci around ancestral- or novel-sex-determining regions could support their role in sex chromosome turnover.
In agreement with this hypothesis, a recent transcriptome analysis in Rumex shows that pollen-biased expression (relative to expression in flower buds or leaves) is enhanced among XY-linked genes compared to autosomal genes or compared to hemizygous genes that are only linked to the X (Sandler et al., 2018, personal communication). 
In addition, Y-linked genes are over-expressed relative to X-linked genes in pollen (but not in flower buds or leaves). 
This suggests that the spread of neo-Y chromosomes in this clade could have been favoured through linkage with haploid selected genes. 
%maybe add a flag here for switching to clade-level expectations:
In general, we predict that haploid selection increases lability of sex-determining systems, particularly because haploid selection can cause transitions that increase or decrease sex-linkage (e.g., the final state of the red line in Fig \ref{fig:SexRatioBad}B is the starting state in Fig \ref{fig:SexRatioBad}A). 
Frequent turnovers driven by haploid selection may help to explain the relative rarity of heteromorphic sex chromosomes in plants, which are thought to experience more selection during their multicellular haploid stage.

>>>>

\end{flushleft}
% Please keep the abstract below 300 words
%currently 308 words
\section*{Abstract}
Sex determination is remarkably dynamic; many taxa display shifts in the location of sex-determining loci or the evolution of entirely new sex-determining systems. 
Predominant theories for why we observe such transitions generally conclude that novel sex-determining systems are favoured by selection if they equalise the sex ratio or increase linkage with a locus that experiences different selection in males vs.\ females. 
We use population genetic models to extend these theories in two ways: 
(1) We explicitly consider selection on loci very tightly linked to the ancestral sex-determining loci, e.g., within the non-recombining region of the ancestral sex chromosomes. 
Variation at such loci can favour the spread of new sex-determining systems in which the heterogametic sex changes (XY to ZW or ZW to XY) and the new sex-determining region is less closely linked (or even unlinked) to the locus under selection, which is not apparent from previous theory. 
(2) We consider selection upon haploid genotypes either during gametic competition (e.g., pollen competition) or meiosis (i.e., non-Mendelian segregation), which can cause the zygotic sex ratio to become biased. 
We find that associations with haploid selected loci can drive transitions between sex-determining systems, without requiring selection to act differently in diploid males vs.\ females. 
Further, with haploid selection, transitions between male and female heterogamety can also evolve where linkage with the sex-determining locus is weakened. 
Unexpectedly, we find that the selective strengths exerted by the sex-ratio and associations with selected alleles are equivalent.
This allows sex-ratio biases to increase or decrease with the spread of new sex chromosomes, and implies that transitions between sex-determining systems cannot be simply predicted by selection to equalise the sex ratio. 
Overall, our models reveal that transitions between sex-determining systems, particularly transitions where the heterogametic sex changes, can be driven by loci in previously unpredicted genomic locations that experience selection during diploid and/or haploid phases.
%Our results predict conditions under which sex-determining systems are likely to be labile and draw connections with sex-ratio evolution.

@@

%Tight linkage Model Feature and Result
One novel feature of the models developed here is that, when considering loci in very tight linkage with the ancestral sex-determining locus (e.g., within the non-recombining region of a sex chromosome), % in the results section: `Tight linkage with the ancestral sex-determining locus'). 
%\textcolor{red}{[vDK2007 does "explicitly" consider this as well (although on pg 15 of their SOM they say they will ignore equilibrium A, as it is stable under a "limited range of biologically relevant parameter combinations"; perhaps haploid selection changes this impression), but only for cis-GSD transitions. They find that if equil B is stable the neo-Y never spreads when r=0. vDK2010 does not "explicitly" calculate pX and pY, which are required to see if their equation 7 is positive or negative, but nowhere do they seem to say that weaker linkage \textit{can't} evolve given r initially small.]}
we explicitly calculate ancestral equilibrium allele frequencies.
This analysis uncovers that, under various forms of selection, the ancestral-X can carry suboptimal alleles for females and/or the ancestral-Y can carry female-beneficial alleles. 
This, in turn, can favour a new ZW sex-determining locus that has weaker linkage with loci under selection, which has not been apparent in previous theory \cite{vanDoorn:2010hu}. 
That is, very tight linkage with the ancestral GSD locus can favour trans-GSD transitions in which associations with selected loci are actually weakened.

@@

Here, we assume that the $\mathbf{M}$ locus is `epistatically dominant' over the $\mathbf{X}$ locus such that zygotes with at least one $m$ allele develop as females with probability $k$ and as males with probability $1-k$, regardless of the $\mathbf{X}$ locus genotype. 
With $k=0$, the $m$ allele is a masculinizer (a neo-Y allele) and with $k=1$ the $m$ allele is a feminizer (a neo-W allele). 
With intermediate $k$, we can interpret $m$ as an environmental sex-determination (ESD) allele, such that zygotes develop as females in a proportion ($k$) of the environments they experience. 
The assumption that derived sex-determining loci are epistatically dominant is motivated by empirical systems in which multiple sex determining alleles segregate (i.e., X, Y, Z, and W alleles present), such as, cichlid fish~\cite{Ser:2010iq}, platyfish (\textit{Xiphophorus maculatus}~\cite{Kallman1968}), houseflies (\textit{Musca domestica}~\cite{Dubendorfer2002}), western clawed frogs (\textit{Xenopus tropicalis}~\cite{Roco2015}) and \textit{Rana rugosa}~\cite{Ogata:2007jm}.  
Nevertheless, our supplementary analysis file (\nameref{file:Mathematica}) allows other dominance relationships between loci to be specified. 
See also the supplementary material in \cite{vanDoorn:2007eu} for a numerical analysis that relaxes the assumption of epistatic dominance.
%We also analyze a model of maternally-controlled environmental sex-determination, where mothers with at least one $m$ allele produce daughters with probability $k$.

@@

\noindent\textit{Conclusion 1A:}
\textbf{selection on loci within the non-recombining region around the sex-determining locus ($r\approx0$) allows the invasion of less closely linked sex-determining alleles ($r<R\leq1/2$) during a trans-GSD transition} (XY $\leftrightarrow$ ZW).
\vspace{0.5cm}

\noindent This does not depend on the form of selection maintaining a polymorphism (sexually-antagonistic selection, overdominance, ploidally-antagonistic selection, or some combination, Fig \ref{fig:SexAntagTighter}, \nameref{fig:temporalOverdominance}, \nameref{fig:regionPloidAntag}, and \nameref{fig:SexAntagTighterMaleDrive}).  
The conditions become more restrictive, however, with increasing recombination ($R$) between the new sex-determining locus and the selected locus.  
The invasion of completely unlinked neo-W alleles ($R = 1/2$) can occur with overdominance in males or with haploid selection (as shown for the case of male drive opposed by diploid selection in \cite{Kozielska:2010vm}) but is not possible with only sexually-antagonistic selection if selection is directional in each diploid sex (see \nameref{file:Mathematica}).
\textcolor{red}{In absence of haploid selection Conclusion 1A can also be reached from Equation 7 in \cite{vanDoorn:2010hu}; for example, with no polymorphism on the Y ($V_Y=0$) and an allelic substitution favoured in females ($\alpha^f,\alpha_X^f>0$) a loosely linked neo-W can invade given the allelic substitution is sufficiently disfavoured on the X in males ($\alpha_X^m < -2\alpha_X^f$), although it is unclear from their implicit equation if and when such an equilibrium is stable.}  
%\vspace{0.5cm}

@@

However, with haploid selection and some ancestral sex-linkage ($r<1/2$; allowing allele frequency differences on the X and Y), the additional term in Eq \eqref{eq:lambda_neoW} can be positive.
This implies the following two conclusions: 
\vspace{0.5cm}

%\noindent\textit{Conclusion 1B:}\
%\textbf{when there is an intermediate rate of recombination between the ancestral sex-determining allele and a selected locus ($0<r<1/2$), haploid selection allows less closely linked sex-determining alleles ($r<R$) to invade during trans-GSD transitions ($\lambda_{W'}^{(XY)}>1$).}
%%\textcolor{red}{[new result because Kozielska had ancestrally-tight sex-linkage and Ubeda had ESD]}
%\vspace{0.5cm}

\noindent\textit{Conclusion 1B:}\
%\textbf{new sex-determining alleles (causing trans-GSD transitions) can spread if they arise in looser linkage with a locus that experiences haploid selection ($r<R$). }
\textbf{haploid selection on loci weakly linked to the sex-determining locus ($0<r<1/2$) allows the invasion of less closely linked sex-determining alleles ($r<R\leq1/2$) during a trans-GSD transition} (XY $\leftrightarrow$ ZW).
\vspace{0.5cm}

\noindent{\textit{Conclusion 2:}\
\textbf{new sex-determining alleles (causing cis- \textit{or} trans-GSD transitions) can spread if they arise in closer linkage with a locus that experiences haploid selection ($R<r$). }
%(This is different but similar to Ubeda et al (2015), who started from ESD)
\vspace{0.5cm}

\noindent The numerical results of \cite{Ubeda:2015fx} are consistent with Conclusion 1B, but arise from an ESD to GSD transition, where the ancestral equilibrium is considerably simpler.
Conclusion 2 echoes the diploid-selection-only results of \cite{vanDoorn:2007eu,vanDoorn:2010hu}, but greatly extends the number of selective scenarios favouring transitions.

%These transitions are unusual because, when $R>r$, associations that selection has built up between alleles more favourable in one sex and alleles that determine sex will be weakened. 
%Mean diploid fitness therefore decreases during heterogametic transitions that create looser sex-linkage (Fig \ref{fig:Combination_MeanFit}B,D). 
%Eq \eqref{eq:lambda_neoW} shows that, with weak selection, neo-W alleles can invade an XY system for a large number of selective regimes. 
%To clarify the parameter space under which $\lambda_{W'}^{(XY)}>1$ , we consider several special cases. 
%Firstly, if the $\mathbf{A}$ locus is unlinked to the ancestral sex-determining region ($r=1/2$), a more closely linked neo-W ($R<1/2$) can always invade because there is no ancestral association between $A$ alleles and sex chromosomes in males, $\left( \hat{p}^\male_Y-\hat{p}^\male_X \right)=0$, see Eq \eqref{eq:freq_diffs}. 
%The second term in Eq \eqref{eq:lambda_neoW} therefore disappears and invasion depends only on the sign of $(r-R)$, as in the case of the neo-Y. 
%Indeed, invasion typically occurs when the neo-W is more closely linked to the selected locus than the ancestral sex-determining region (Fig \ref{fig:Combination_Centimorgans}).
To clarify the parameter space under which invasion occurs despite looser sex-linkage ($\lambda_{W'}^{(XY)}>1$ despite $R>r$; Conclusion 1B), we focus on the special case where $R=1/2$ and $r<1/2$ (e.g., the selected locus is on the ancestral sex chromosome and the novel sex-determining locus arises on an autosome). 
In Table \ref{tab:specialcases} we give the conditions where invasion occurs when we further assume that haploid selection only occurs in one sex (e.g., during male meiosis only) and dominance coefficients are equal in the two sexes, $h^\female=h^\male$. 
These special cases indicate that neo-W invasion occurs for a large fraction of the parameter space, even if the neo-W uncouples the sex-determining locus from a locus under selection. 
We show that neo-W alleles can spread for when they are more closely \textit{or} more loosely linked to a locus that experiences haploid selection in Fig \ref{fig:Combination_Centimorgans} (c.f., Fig \ref{fig:SexAntagTighter}A for diploid sexually-antagonistic selection alone). 
In Fig \ref{fig:SexRatioBad}, we also show that trans-GSD transitions in sex determination occur during which linkage with a haploid selected locus increases \textit{or} decreases.

@@

Previous research has shown that when the ancestral sex-determining locus is linked to a locus that experiences haploid selection (e.g., meiotic drive), alleles at a new, unlinked sex-determining locus can invade and restore equal sex ratios~\cite{Kozielska:2010vm}. 
Consider, for example, the case where the $\mathbf{A}$ locus is linked to the ancestral sex-determining locus ($r<1/2$) and experiences meiotic drive in males only ($\alpha^\male_\Delta \neq 0$, $\alpha^\female_\Delta=0$), without gametic competition ($t^\female=t^\male=0$).
In this case, the zygotic sex ratio can be initially biased only if the ancestral sex-determining system is XY (Fig \ref{fig:model_outline}A and Fig \ref{fig:SexRatioBad}B). 
%If the ancxestral sex-determining system is ZW, the zygotic sex ratio will be 1:1 because diploid sex is determined by the proportion of Z-bearing versus W-bearing eggs and meiosis in females is fair (Fig \ref{fig:Combination_Turnover}D).
If Fisherian sex-ratio selection were dominant, we would thus expect a difference in the potential for XY to ZW and ZW to XY transitions. 
%However, to leading order with selection weak relative to recombination, we find that sex-ratio selection favours the spread of a neo-W by an amount equl to the spread of a neo-Y into a ZW system (relabeling XY as the heterogametic sex and reversing the labels of males and females in Table \ref{tab:specialcases})
However, invasion by a neo-W allele into an XY system and invasion by a neo-Y allele into a ZW system occur under the same conditions ($\lambda_{Y'}^{(XY)} = \lambda_{W'}^{(ZW)}$ and $\lambda_{W'}^{(XY)}=\lambda_{Y'}^{(ZW)}$, at least to order $\epsilon^2$), implying that, surprisingly,
\vspace{0.5cm}

@@

%\noindent\textcolor{red}{\textit{Conclusion 3:}
%\textbf{when selection is weak relative to recombination, sex-ratio selection is exactly equal in magnitude to the selection imposed by associations with selected alleles. } }
%%\textcolor{red}{[new result]}
%\vspace{0.5cm}

@@

We next consider the case where the new sex-determining mutation, $m$, causes sex to be determined probabilistically or by heterogeneous environmental conditions (environmental sex determination, ESD), with individuals carrying allele $m$ developing as females with probability $k$.
%\textcolor{red}{In our deterministic model this means the fraction female is exactly $k$, even when $m$ is rare (exploring the effect of the variance induced by ESD would also be interesting).}
In our deterministic model this means the fraction female is exactly $k$, even when $m$ is rare (exploring the effect of the variance induced by ESD would also be interesting).
Here, we do not assume that the environmental conditions that determine sex also differentially affect the fitness of males versus females. 
Such correlations can favour environmental sex-determining systems by allowing each sex to be produced in the environment in which it has highest fitness; in the absence of these correlations, previous theory would predict that ESD is favoured when it produces more equal sex ratios than the ancestral system (see reviews by~\cite{Charnov:1982wg,Bull:1983vi,West:2009we}). %(e.g., maternal condition, mate quality, age, or host size)

@@

\noindent\textit{Conclusion 4:}
\textbf{
transitions from genetic to environmental sex-determination are not straightforwardly predicted by selection to balance the zygotic sex ratio when haploid selection is present. 
}
%\textcolor{red}{[new result]}

@@

%HAPLOID SELECTION (genomic location of haploid selected loci and comparative patterns) 
We have shown that the spread of new sex-determining systems can be driven by loci experiencing haploid selection. 
%Our results suggest that haploid selected loci that have been involved in driving turnovers between sex-determining systems could occur around the ancestral- or novel-sex-determining regions. 
%Our results suggest that polymorphic loci that are ancestrally sex-linked and under sex-specific selection could also drive heterogametic transitions between sex-determining systems. 
%As with sexually-antagonistic selection, the presence of haploid selected loci around ancestral- or novel-sex-determining regions could support their role in sex chromosome turnover.
In agreement with this hypothesis, a recent transcriptome analysis in Rumex shows that pollen-biased expression (relative to expression in flower buds or leaves) is enhanced among XY-linked genes compared to autosomal genes or compared to hemizygous genes that are only linked to the X 
\cite{Sandler264382}. 
In addition, Y-linked genes are over-expressed relative to X-linked genes in pollen (but not in flower buds or leaves). 
This suggests that the spread of neo-Y chromosomes in this clade could have been favoured through linkage with haploid selected genes. 
%maybe add a flag here for switching to clade-level expectations:
In general, we predict that haploid selection increases lability of sex-determining systems, particularly because haploid selection can cause transitions that increase or decrease sex-linkage (e.g., the final state of the red line in Fig \ref{fig:SexRatioBad}B is the starting state in Fig \ref{fig:SexRatioBad}A). 
Frequent turnovers driven by haploid selection may help to explain the relative rarity of heteromorphic sex chromosomes in plants, which are thought to experience more selection during their multicellular haploid stage.
